---
title: "MLDataR_tutorial_YT_Hutson"
author: "ASG"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries
```{r}
library(MLDataR)
library(ConfusionTableR)
library(dplyr)
library(tidymodels)
```

# pre-processing data
## thyroid_df
Creating a dataframe of the thyroid disease dataset

```{r}
thyroid_df <- MLDataR::thyroid_disease
```


The aim is to train a supervised learning model to predict whether the patient is sick or not. So, we convert the outcome variable 'ThryroidClass' into a factor. We also remove the uninformative variable ref_src. Note "ThryroidClass" incorrect spelling of thyroid in original data set.

```{r}
thyroid_df$ThryroidClass <- as.factor(thyroid_df$ThryroidClass)
thyroid_df <- thyroid_df %>% select(-ref_src)
```

## selecting for patients with all variables reported
```{r}
thyroid_df <- thyroid_df[complete.cases(thyroid_df), ] # selecting complete rows; see documentation for complete.cases function
```

## Splitting the data into training and testing datasets
```{r}
set.seed(1)
split_data <- 
  rsample::initial_split(
  thyroid_df,prop = 2/3
  )
thyroid_training_set <- training(split_data)
thyroid_test_set <- testing(split_data)
```

Setting a seed makes analysis reproducible by generating the same splits every time.

### accessing the training and the test data sets
```{r}
glimpse(thyroid_training_set)
glimpse(thyroid_test_set)
```

To note: all variables are numeric (dbl) except the ThryroidClass variable which is a factor. This is useful and we can normalize the numeric predictors to a mean of zero and an SD of one. This eliminates outsize influence of outliers.

## Specifying a recipe with predictors, outcomes and pre-processing steps
```{r}
set.seed(1)
thyroid_recipe <- 
  recipe(thyroid_training_set,ThryroidClass ~ .) %>%
  step_normalize(all_predictors()) %>% 
  step_zv(all_predictors())
```
 Checking the recipe
```{r}
thyroid_recipe
```
 
## specifying models & engines

```{r}
set.seed(1)
logistic_model <- parsnip::logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

# specifying a second model
set.seed(1)
rand_forest_model <- parsnip::rand_forest() %>% 
  set_engine("ranger") %>% 
  # ensure ranger package is installed but do not "library' in
  set_mode("classification")
```

## package the recipe into workflows
```{r}
logistic_workflow <- 
  workflow() %>% 
  add_recipe(thyroid_recipe) %>% 
  add_model(logistic_model)

# another workflow for the random forest model
rand_forest_workflow <- 
  workflow() %>% 
  add_recipe(thyroid_recipe) %>% 
  add_model(rand_forest_model)
```

```{r}
# checking the workflows
logistic_workflow
rand_forest_workflow
```

## fitting the models

```{r}
# fit the model to the training data
thyroid_log_fit <- 
  logistic_workflow %>% 
  fit(data = thyroid_training_set)

thyroid_randF_fit <- 
  rand_forest_workflow %>% 
  fit(data = thyroid_training_set)
```

```{r}
# extract the fit parameters
log_fitd <- 
  thyroid_log_fit %>% 
  extract_fit_parsnip()

randF_fitd <- 
  thyroid_randF_fit %>% 
  extract_fit_parsnip()
```

## fit test data and generate predictions

```{r}
thyroid_log_preds <- 
  thyroid_log_fit %>% 
  predict(thyroid_test_set) %>% 
  bind_cols(thyroid_test_set)

thyroid_randF_preds <- 
  thyroid_randF_fit %>% 
  predict(thyroid_test_set) %>% 
  bind_cols(thyroid_test_set)
```

```{r} 
# convret the predictions to factors
thyroid_log_preds$ThryroidClass <- as.factor(thyroid_log_preds$ThryroidClass)

thyroid_randF_preds$ThryroidClass <- 
  as.factor(thyroid_randF_preds$ThryroidClass)
```

```{r}
# evaluate the goodness of fits as a confusion matrix
log_confusion_matrix <- 
  ConfusionTableR::binary_class_cm(
    train_labels = thyroid_log_preds$ThryroidClass,
    truth_labels = thyroid_log_preds$.pred_class
  )

randF_confusion_matrix <- 
  ConfusionTableR::binary_class_cm(
    train_labels = thyroid_randF_preds$ThryroidClass,
    truth_labels = thyroid_randF_preds$.pred_class
  )
```
```{r}
# view the goodness of fits
log_confusion_matrix$confusion_matrix
log_confusion_matrix$record_level_cm
randF_confusion_matrix$confusion_matrix
randF_confusion_matrix$record_level_cm
```

